Indicators for Ecosystem Condition in Norway

### 2023-10-05 NDVI_natopen.Rmd

@@

**How does NDVI vary over the years?**
```{r NDVI-values-over-time-ndvin, fig.cap="A figure showing NDVI values for naturally open ecosystem over time for six different main nature types in varying degrees of condition.}
SentinelNDVI.natopen %>%
  filter(tilstand!="") %>%
  ggplot(aes(x = year, y = mean)) +
  geom_point() +
  facet_grid(tilstand ~ hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2, size = 12)) +
  ggtitle("NDVI across years for different ecosystem types in all condition classes") +
  labs(y = "NDVI values (Sentinel-2)", x = "Year")
```

@@

```{r NDVI-basic-ndvin, warning=FALSE, fig.cap="NDVI values in basic natopen ecosystem types in good condition.}
SentinelNDVI.natopen %>%
  filter(tilstand == "God") %>%
  filter(year == kartleggingsaar) %>%
  ggplot(aes(x = subtype, y = mean)) +
  geom_violin() +
  geom_point(size = 0.7, shape = 16, color = "grey", position = position_jitter(.1) ) +
  facet_wrap(~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2, size = 12)) +
  ggtitle("NDVI for different ecosystem sub-types in good condition") +
  labs(y = "NDVI values (Sentinel-2)", x = "Ecosystem sub-type")

@@

**How does NDVI vary across regions in natopens (major types) in good condition? (using only NDVI data matching NiN-mapping years)**
```{r ndvi-regions-ndvin, warning=FALSE, fig.cap="DVI variation across regions in natopen ecosystems (major types) in good condition.}
SentinelNDVI.natopen %>%
  filter(tilstand == "God") %>%
  filter(year == kartleggingsaar) %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = region, y = mean)) +
  geom_violin() +
  geom_point(size = 0.7, shape = 16, color = "grey", position = position_jitter(.1) ) +
  facet_wrap(~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2, size = 9)) +
  ggtitle("NDVI for different ecosystems in good condition across regions") +
  labs(y = "NDVI values (Sentinel-2)", x = "Region")
```

@@

**How does NDVI vary across condition classes (for major types and NDVI data matching NiN-mapping years)?**
```{r ndvi-condition-ndvin, warning=FALSE, fig.cap="NDVI values across natopen sites in different ecological condition.}
SentinelNDVI.natopen %>%
  filter(year == kartleggingsaar) %>%
  filter(tilstand != "") %>%
  ggplot(aes(x = tilstand, y = mean)) +
  geom_violin() +
  geom_point(size = 0.7, shape = 16, color = "grey", position = position_jitter(.1) ) +
  facet_wrap(~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2)) +
  ggtitle("NDVI for different ecosystems in different condition classes") +
  labs(y = "NDVI values (Sentinel-2)", x = "Condition")
```
The plots suggest a slight decrease in NDVI as condition deteriorates. 
Overlaps between condition classes are generally very large.

>>>>

# (PART\*) R & D{-}

@@

**How does NDVI vary over the years?**
```{r NDVI-values-over-time-ndvin, fig.cap="A figure showing NDVI values for naturally open ecosystem over time for six different main nature types in varying degrees of condition."}
SentinelNDVI.natopen %>%
  filter(tilstand!="") %>%
  ggplot(aes(x = year, y = mean)) +
  geom_point() +
  facet_grid(tilstand ~ hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2, size = 12)) +
  ggtitle("NDVI across years for different ecosystem types in all condition classes") +
  labs(y = "NDVI values (Sentinel-2)", x = "Year")
```

@@

```{r NDVI-basic-ndvin, warning=FALSE, fig.cap="NDVI values in basic natopen ecosystem types in good condition."}
SentinelNDVI.natopen %>%
  filter(tilstand == "God") %>%
  filter(year == kartleggingsaar) %>%
  ggplot(aes(x = subtype, y = mean)) +
  geom_violin() +
  geom_point(size = 0.7, shape = 16, color = "grey", position = position_jitter(.1) ) +
  facet_wrap(~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2, size = 12)) +
  ggtitle("NDVI for different ecosystem sub-types in good condition") +
  labs(y = "NDVI values (Sentinel-2)", x = "Ecosystem sub-type")

@@

**How does NDVI vary across regions in natopens (major types) in good condition? (using only NDVI data matching NiN-mapping years)**
```{r ndvi-regions-ndvin, warning=FALSE, fig.cap="NDVI variation across regions in natopen ecosystems (major types) in good condition."}
SentinelNDVI.natopen %>%
  filter(tilstand == "God") %>%
  filter(year == kartleggingsaar) %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = region, y = mean)) +
  geom_violin() +
  geom_point(size = 0.7, shape = 16, color = "grey", position = position_jitter(.1) ) +
  facet_wrap(~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2, size = 9)) +
  ggtitle("NDVI for different ecosystems in good condition across regions") +
  labs(y = "NDVI values (Sentinel-2)", x = "Region")
```

@@

**How does NDVI vary across condition classes (for major types and NDVI data matching NiN-mapping years)?**
```{r ndvi-condition-ndvin, warning=FALSE, fig.cap="NDVI values across natopen sites in different ecological condition."}
SentinelNDVI.natopen %>%
  filter(year == kartleggingsaar) %>%
  filter(tilstand != "") %>%
  ggplot(aes(x = tilstand, y = mean)) +
  geom_violin() +
  geom_point(size = 0.7, shape = 16, color = "grey", position = position_jitter(.1) ) +
  facet_wrap(~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2)) +
  ggtitle("NDVI for different ecosystems in different condition classes") +
  labs(y = "NDVI values (Sentinel-2)", x = "Condition")
```
The plots suggest a slight decrease in NDVI as condition deteriorates. 
Overlaps between condition classes are generally very large.

### 2023-10-05 NDVI_wetland.Rmd

**How does NDVI vary across regions in wetlands (major types) in good condition? (using only NDVI data matching NiN-mapping years)**
```{r ndvi-regions-ndviw, warning=FALSE, fig.cap="DVI variation across regions in wetland ecosystems (major types) in good condition. V1 = minerotrophic mires, V3 = ombrotrophic bogs, V4 = cold springs, V9 = semi-natural mires, V10 = semi-natural wet meadows."}
SentinelNDVI.wetland %>%
  filter(tilstand == "God") %>%
  filter(year == kartleggingsaar) %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = region, y = mean)) +
  geom_violin() +
  geom_point(size = 0.7, shape = 16, color = "grey", position = position_jitter(.1) ) +
  facet_wrap(~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2, size = 9)) +
  ggtitle("NDVI for different ecosystems in good condition across regions") +
  labs(y = "NDVI values (Sentinel-2)", x = "Region")
```
There is no obvious trends in geography, with the exception of V3 which shows lower values in Central Norway.

>>>>

**How does NDVI vary across regions in wetlands (major types) in good condition? (using only NDVI data matching NiN-mapping years)**
```{r ndvi-regions-ndviw, warning=FALSE, fig.cap="NDVI variation across regions in wetland ecosystems (major types) in good condition. V1 = minerotrophic mires, V3 = ombrotrophic bogs, V4 = cold springs, V9 = semi-natural mires, V10 = semi-natural wet meadows."}
SentinelNDVI.wetland %>%
  filter(tilstand == "God") %>%
  filter(year == kartleggingsaar) %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = region, y = mean)) +
  geom_violin() +
  geom_point(size = 0.7, shape = 16, color = "grey", position = position_jitter(.1) ) +
  facet_wrap(~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0.2, size = 9)) +
  ggtitle("NDVI for different ecosystems in good condition across regions") +
  labs(y = "NDVI values (Sentinel-2)", x = "Region")
```
There is no obvious trends in geography, with the exception of V3 which shows lower values in Central Norway.
